---
title: "Strompreise"
output:
  html_notebook: default
  pdf_document: default
---
Load the libraries.
```{r message=FALSE}
library(RSNNS)
RSNNS::getSnnsRFunctionTable()
```

Read the data and add previous Strompreis.
```{r}
strompreise <- read.csv("/home/david/work/r/strom/psbi/Daten/Strompreise.csv")
prevPreis = strompreise$Strompreis[-8784]
pprevPreis = prevPreis[-8783]
ppprevPreis = pprevPreis[-8782]
prevPreis = prevPreis[-1]
prevPreis = prevPreis[-1]
pprevPreis = pprevPreis[-1]
#prevPreis = prevPreis[-8783]
#pprevPreis = strompreise$Strompreis[-]
strompreise = strompreise[-1,]
strompreise = strompreise[-1,]
strompreise = strompreise[-1,]
strompreise$prevPreis <- prevPreis
strompreise$pprevPreis <- pprevPreis
strompreise$ppprevPreis <- ppprevPreis
strompreise

```

Normalize the data using the RSNNS library.

Definitely omit column DatumUhrzeit as it is a redundant, combined, badly-analyzable value of the other columns.

Consider omitting column Jahr as it is the same for every value.

Wochentag is first mapped to values 1-7 using as.numeric and then normalized by RSNNS. The fact, that we can't control which day 1 will be, has no effect, as the denormalization will result in the corresponding day again.

We should optimally convert the data frame to a matrix, normalize it and convert it back to a data frame. But I was not able to do so, so this is the workaround.
```{r}
# Normalize every column.
normalized = NULL
normalized$Jahresstunde <- normalizeData(strompreise$Jahresstunde)
# n$Jahr <- normalizeDate(s$Jahr)
normalized$Monat <- normalizeData(strompreise$Monat, type = "0_1")
normalized$Tag <- normalizeData(strompreise$Tag, type = "0_1")
normalized$Stunde <- normalizeData(strompreise$Stunde, type = "0_1")
normalized$Woche <- normalizeData(strompreise$Woche, type = "0_1")
normalized$Wochentag <- normalizeData(as.numeric(strompreise$Wochentag), type = "0_1")
normalized$Temperatur <- normalizeData(strompreise$Temperatur, type = "0_1")
normalized$Bedeckung <- normalizeData(strompreise$Bedeckung, type = "0_1")
normalized$Windgeschwindigkeit <- normalizeData(strompreise$Windgeschwindigkeit, type = "0_1")
normalized$Globalstrahlung <- normalizeData(strompreise$Globalstrahlung, type = "0_1")
normalized$Strompreis <- normalizeData(strompreise$Strompreis, type = "0_1")
normalized$prevPreis <- normalizeData(strompreise$prevPreis, type = "0_1")
normalized$pprevPreis <- normalizeData(strompreise$pprevPreis, type = "0_1")
normalized$ppprevPreis <- normalizeData(strompreise$ppprevPreis, type = "0_1")
# Bind columns to matrix.
converted <- cbind(#normalized$Jahresstunde, 
                   #normalized$Monat,
                   #normalized$Tag,
                   #normalized$Stunde,
                   #normalized$Woche,
                   #normalized$Wochentag,
                   normalized$Temperatur,
                   normalized$Bedeckung,
                   normalized$Windgeschwindigkeit,
                   normalized$Globalstrahlung,
                   normalized$Strompreis,
                   normalized$prevPreis,
                   normalized$pprevPreis,
                   normalized$ppprevPreis)
# Covert matrix to data frame.
dataframe <- as.data.frame(converted)
# Rename columns.
#colnames(dataframe) <- c("Monat", "Tag", "Stunde", "Woche", "Wochentag", "Temperatur", "Bedeckung", "Windgeschwindigkeit", "Globalstrahlung", "Strompreis", "prevPreis", "pprevPreis")
colnames(dataframe) <- c("Temperatur", "Bedeckung", "Windgeschwindigkeit", "Globalstrahlung","Strompreis", "prevPreis", "pprevPreis", "ppprevPreis")
```

Have a look at the dataframe.
```{r}
str(dataframe)
```

Split data to inputs and target.
```{r}
inputs <- dataframe
inputs$Strompreis <- NULL
targets <- dataframe$Strompreis
```

Split data for training and test.
```{r}
splitdata <- splitForTrainingAndTest(inputs, targets, ratio=0.5)
```

Fit the model. mlp allows to put test data in as well and will test the mlp itself. 

Configure the size at your choice.
```{r}
model <- mlp(splitdata$inputsTrain,
             splitdata$targetsTrain,
             size=c(8,9,8,7),
             maxit=3000,
             #hiddenActFunc = "Act_Elliott",
             inputsTest=splitdata$inputsTest,
             targetsTest=splitdata$targetsTest)
model
plotIterativeError(model)
#plotRegressionError(model)
```

Denormalize the column Strompreis of the data frame utilizing the norm parameters of the large list using RSNNS.
```{r}
prediction <- denormalizeData(model$fittedTestValues,
                              getNormParameters(normalized$Strompreis))
realTest <- denormalizeData(splitdata$targetsTest,
                            getNormParameters(normalized$Strompreis))
```

Analyse the result.
```{r}
result <- cbind(prediction,realTest)
result
plot(result)
cor.test(prediction, realTest, method="pearson")
```

Clean up.
```{r message=FALSE}
remove(strompreise)
remove(normalized)
remove(converted)
remove(dataframe)
remove(inputs)
remove(splitdata)
remove(model)
remove(prediction)
remove(realTest)
remove(targets)
remove(result)
```