---
title: "Strompreise"
output:
  html_notebook: default
  pdf_document: default
---
# Strompreise
## Libraries
```{r message=FALSE}
library(RSNNS)
```

## File read, preprocess data
```{r}
raw <- read.csv("Kopie von Wetterdaten_Strompreise_mehrere_Standorte.csv")
strompreise <- raw
prevPreis = strompreise$Strompreis[-8784]
pprevPreis = prevPreis[-8783]
ppprevPreis = pprevPreis[-8782]
prevPreis = prevPreis[-1]
prevPreis = prevPreis[-1]
pprevPreis = pprevPreis[-1]
#prevPreis = prevPreis[-8783]
#pprevPreis = strompreise$Strompreis[-]
strompreise = strompreise[-1,]
strompreise = strompreise[-1,]
strompreise = strompreise[-1,]
strompreise$prevPreis <- prevPreis
strompreise$pprevPreis <- pprevPreis
strompreise$ppprevPreis <- ppprevPreis
```
## Normalization

Normalize the data using the RSNNS library.

Definitely omit column DatumUhrzeit as it is a redundant, combined, badly-analyzable value of the other columns.

Consider omitting column Jahr as it is the same for every value.

Wochentag is first mapped to values 1-7 using as.numeric and then normalized by RSNNS. The fact, that we can't control which day 1 will be, has no effect, as the denormalization will result in the corresponding day again.

We should optimally convert the data frame to a matrix, normalize it and convert it back to a data frame. But I was not able to do so, so this is the workaround.
```{r}
# Normalize every column.
normalized = NULL
normalized$Jahresstunde <- normalizeData(strompreise$Jahresstunde)
normalized$Jahr <- normalizeData(strompreise$Jahr)
normalized$Monat <- normalizeData(strompreise$Monat, type = "0_1")
normalized$Tag <- normalizeData(strompreise$Tag, type = "0_1")
normalized$Stunde <- normalizeData(strompreise$Stunde, type = "0_1")
# normalized$Woche <- normalizeData(strompreise$Woche, type = "0_1")
# normalized$Wochentag <- normalizeData(as.numeric(strompreise$Wochentag), type = "0_1")
normalized$D_Temperatur <- normalizeData(strompreise$D_Temperatur, type = "0_1")
normalized$M_Temperatur <- normalizeData(strompreise$M_Temperatur, type = "0_1")
normalized$S_Temperatur <- normalizeData(strompreise$S_Temperatur, type = "0_1")
normalized$D_Bedeckung <- normalizeData(strompreise$D_Bedeckung, type = "0_1")
normalized$M_Bedeckung <- normalizeData(strompreise$M_Bedeckung, type = "0_1")
normalized$S_Bedeckung <- normalizeData(strompreise$S_Bedeckung, type = "0_1")
normalized$D_Windgeschwindigkeit <- normalizeData(strompreise$D_Windgeschwindigkeit, type = "0_1")
normalized$M_Windgeschwindigkeit <- normalizeData(strompreise$M_Windgeschwindigkeit, type = "0_1")
normalized$S_Windgeschwindigkeit <- normalizeData(strompreise$S_Windgeschwindigkeit, type = "0_1")
normalized$Globalstrahlung <- normalizeData(strompreise$Globalstrahlung, type = "0_1")
normalized$Strompreis <- normalizeData(strompreise$Strompreis, type = "0_1")
normalized$prevPreis <- normalizeData(strompreise$prevPreis, type = "0_1")
normalized$pprevPreis <- normalizeData(strompreise$pprevPreis, type = "0_1")
normalized$ppprevPreis <- normalizeData(strompreise$ppprevPreis, type = "0_1")
# Bind columns to matrix.
converted <- cbind(#normalized$Jahresstunde, 
                   #normalized$Monat,
                   #normalized$Tag,
                   #normalized$Stunde,
                   #normalized$Woche,
                   #normalized$Wochentag,
                   normalized$D_Temperatur,
                   normalized$M_Temperatur,
                   normalized$S_Temperatur,
                   normalized$D_Bedeckung,
                   normalized$M_Bedeckung,
                   normalized$S_Bedeckung,
                   normalized$D_Windgeschwindigkeit,
                   normalized$M_Windgeschwindigkeit,
                   normalized$S_Windgeschwindigkeit,
                   normalized$Globalstrahlung,
                   normalized$Strompreis,
                   normalized$prevPreis,
                   normalized$pprevPreis,
                   normalized$ppprevPreis)
# Covert matrix to data frame.
dataframe <- as.data.frame(converted)
# Rename columns.
#colnames(dataframe) <- c("Monat", "Tag", "Stunde", "Woche", "Wochentag", "Temperatur", "Bedeckung", "Windgeschwindigkeit", "Globalstrahlung", "Strompreis", "prevPreis", "pprevPreis")
colnames(dataframe) <- c("D_Temperatur", "M_Temperatur", "S_Temperatur", "D_Bedeckung", "M_Bedeckung", "S_Bedeckung", "D_Windgeschwindigkeit", "M_Windgeschwindigkeit", "S_Windgeschwindigkeit", "Globalstrahlung","Strompreis", "prevPreis", "pprevPreis", "ppprevPreis")

inputs <- dataframe
inputs$Strompreis <- NULL
targets <- dataframe$Strompreis

splitdata <- splitForTrainingAndTest(inputs, targets, ratio=0.5)
```

## Model and analysis
Fit the model. mlp allows to put test data in as well and will test the mlp itself. 

Configure the size at your choice.
```{r}
model <- mlp(splitdata$inputsTrain,
             splitdata$targetsTrain,
             size=c(20),
             maxit=100,
             #hiddenActFunc = "Act_Elliott",
             inputsTest=splitdata$inputsTest,
             targetsTest=splitdata$targetsTest)
model
plotIterativeError(model)
#plotRegressionError(model)
```

## Prediction
## 1h
Taken from prefitted data.
```{r}
prediction <- denormalizeData(model$fittedTestValues,
                              getNormParameters(normalized$Strompreis))
realTest <- denormalizeData(splitdata$targetsTest,
                            getNormParameters(normalized$Strompreis))
result <- cbind(prediction,realTest)
resultdiff <- prediction - realTest
result
plot(result)
plot(resultdiff)

sum((realTest - prediction)^2) / length(realTest)
```

## Continous 24h
```{r}
continous24 <- as.data.frame(splitdata$inputsTest)
continous24result <- NULL
continous24result$real <- denormalizeData(splitdata$targetsTest,
                                    getNormParameters(normalized$Strompreis))
for(i in 1:24)
{
  raw_i <- paste("raw", i)
  pred_i <- paste("pred", i)
  result_i <- paste("result", i)
  diff_i <- paste("diff", i)
  abs_i <- paste("abs", i)
  sum_i <- paste("sum", i)
  continous24result[[pred_i]] <- denormalizeData(predict(model, continous24),
                                        getNormParameters(normalized$Strompreis)) 
  continous24result[[result_i]] <- cbind(continous24result[[pred_i]], continous24result$real)
  continous24result[[diff_i]] <- continous24result$real - continous24result[[pred_i]]
  continous24result[[abs_i]] <- abs(continous24result[[diff_i]])
  continous24result[[sum_i]] <- sum(continous24result[[diff_i]]^2) / length(continous24result[[diff_i]])
  continous24result[[result_i]]
  #plot(continous24result[[pred_i]])
  plot(continous24result[[result_i]])
  #plot(continous24result[[diff_i]])
  #plot(continous24result[[abs_i]])
  
  # Shift previous prices.
  continous24$ppprevPreis <- continous24$pprevPreis
  continous24$pprevPreis <- continous24$prevPreis
  # IMPORTANT: denormalize output and normalize input, as different paramters might be used.
  continous24$prevPreis <- normalizeData(continous24result[[pred_i]], getNormParameters(normalized$prevPreis))
}
```
## Single 24h
```{r}
output24 <- NULL
input24 <- splitdata$inputsTest

start <- 0
end <- 120
input24 <- splitdata$inputsTest[start:(end+3),]
for(i in 1:(end-start))
{
  output <- predict(model, input24[i,])
  output24[i-start] <- output
  input24[i+1,"prevPreis"] <- output
  input24[i+2,"pprevPreis"] <- output
  input24[i+3,"ppprevPreis"] <- output
}

prediction24 <- denormalizeData(output24,
                                getNormParameters(normalized$Strompreis))

realTest24 <- denormalizeData(splitdata$targetsTest[1:(end-start)],
                              getNormParameters(normalized$Strompreis)) 
result24 <- cbind(prediction24,realTest24)
result24diff <- realTest24 - prediction24
result24diffabs <- abs(result24diff)
result24
plot(result24)
plot(result24diff)
plot(result24diffabs)

sum((realTest24 - prediction24)^2) / length(prediction24)
```

Clean up.
```{r message=FALSE}
remove(strompreise)
remove(normalized)
remove(converted)
remove(dataframe)
remove(inputs)
remove(splitdata)
remove(model)
remove(prediction)
remove(realTest)
remove(targets)
remove(result)
```