---
title: "Strompreise"
output:
  html_notebook: default
  pdf_document: default
---
# Strompreise
## Libraries
```{r message=FALSE}
library(RSNNS)
```

## File read, preprocess data
```{r}
raw <- read.csv("/home/david/Kopie von Wetterdaten_Strompreise_mehrere_Standorte.csv")
strompreise <- raw
strompreise
len = nrow(strompreise)
len
i = 24
namePreis = "revPreis"
nameGlobalstrahlung = "revGlobalstrahlung"
nameDTemp = "revDTemp"
nameDBed = "revDBed"
nameDWind = "revDWind"
nameSTemp = "revSTemp"
nameSBed = "revSBed"
nameSWind = "revSWind"
nameMTemp = "revMTemp"
nameMBed = "revMBed"
nameMWind = "revMWind"
preis = strompreise$Strompreis
globalstrahlung = strompreise$Globalstrahlung
dtemp = strompreise$D_Temperatur
dbed = strompreise$D_Bedeckung
dwind = strompreise$D_Windgeschwindigkeit
stemp = strompreise$S_Temperatur
sbed = strompreise$S_Bedeckung
swind = strompreise$S_Windgeschwindigkeit
mtemp = strompreise$M_Temperatur
mbed = strompreise$M_Bedeckung
mwind = strompreise$M_Windgeschwindigkeit
for(hour in 0:i){
  namePreis = paste("p", namePreis, sep = "")
  nameGlobalstrahlung = paste("p", nameGlobalstrahlung, sep = "")
  nameDTemp = paste("p", nameDTemp, sep = "")
  nameDBed = paste("p", nameDBed, sep = "")
  nameDWind = paste("p", nameDWind, sep = "")
  nameSTemp = paste("p", nameSTemp, sep = "")
  nameSBed = paste("p", nameSBed, sep = "")
  nameSWind = paste("p", nameSWind, sep = "")
  nameMTemp = paste("p", nameMTemp, sep = "")
  nameMBed = paste("p", nameMBed, sep = "")
  nameMWind = paste("p", nameMWind, sep = "")
  len = nrow(strompreise)
  preis = preis[-len]
  globalstrahlung = globalstrahlung[-len]
  dtemp = dtemp[-len]
  dbed = dbed[-len]
  dwind = dwind[-len]
  stemp = stemp[-len]
  sbed = sbed[-len]
  swind = swind[-len]
  mtemp = mtemp[-len]
  mbed = mbed[-len]
  mwind = mwind[-len]
  strompreise = strompreise[-1:-1,]
  strompreise[namePreis] <- preis
  strompreise[nameGlobalstrahlung] <- globalstrahlung
  strompreise[nameDTemp] <- dtemp
  strompreise[nameDBed] <- dbed
  strompreise[nameDWind] <- dwind
  strompreise[nameSTemp] <- stemp
  strompreise[nameSBed] <- sbed
  strompreise[nameSWind] <- swind
  strompreise[nameMTemp] <- mtemp
  strompreise[nameMBed] <- mbed
  strompreise[nameMWind] <- mwind
}
strompreise
```
## Normalization

Normalize the data using the RSNNS library.

Definitely omit column DatumUhrzeit as it is a redundant, combined, badly-analyzable value of the other columns.

Consider omitting column Jahr as it is the same for every value.

Wochentag is first mapped to values 1-7 using as.numeric and then normalized by RSNNS. The fact, that we can't control which day 1 will be, has no effect, as the denormalization will result in the corresponding day again.

We should optimally convert the data frame to a matrix, normalize it and convert it back to a data frame. But I was not able to do so, so this is the workaround.
```{r}
# Normalize every column.
normalized = NULL
#normalized$Jahresstunde <- normalizeData(strompreise$Jahresstunde)
#normalized$Jahr <- normalizeData(strompreise$Jahr)
#normalized$Monat <- normalizeData(strompreise$Monat, type = "0_1")
#normalized$Tag <- normalizeData(strompreise$Tag, type = "0_1")
normalized$Stunde <- normalizeData(strompreise$Stunde, type = "0_1")
# normalized$Woche <- normalizeData(strompreise$Woche, type = "0_1")
#normalized$Wochentag <- normalizeData(strompreise$Wochentag, type = "0_1")
normalized$D_Temperatur <- normalizeData(strompreise$D_Temperatur, type = "0_1")
normalized$M_Temperatur <- normalizeData(strompreise$M_Temperatur, type = "0_1")
normalized$S_Temperatur <- normalizeData(strompreise$S_Temperatur, type = "0_1")
normalized$D_Bedeckung <- normalizeData(strompreise$D_Bedeckung, type = "0_1")
normalized$M_Bedeckung <- normalizeData(strompreise$M_Bedeckung, type = "0_1")
normalized$S_Bedeckung <- normalizeData(strompreise$S_Bedeckung, type = "0_1")
normalized$D_Windgeschwindigkeit <- normalizeData(strompreise$D_Windgeschwindigkeit, type = "0_1")
normalized$M_Windgeschwindigkeit <- normalizeData(strompreise$M_Windgeschwindigkeit, type = "0_1")
normalized$S_Windgeschwindigkeit <- normalizeData(strompreise$S_Windgeschwindigkeit, type = "0_1")
normalized$Globalstrahlung <- normalizeData(strompreise$Globalstrahlung, type = "0_1")
normalized$Strompreis <- normalizeData(strompreise$Strompreis, type = "0_1")
# Bind columns to matrix.
converted <- cbind(#normalized$Jahresstunde, 
                   #normalized$Monat,
                   #normalized$Tag,
                   normalized$Stunde,
                   #normalized$Woche,
                   #normalized$Wochentag,
                   normalized$D_Temperatur,
                   normalized$M_Temperatur,
                   normalized$S_Temperatur,
                   normalized$D_Bedeckung,
                   normalized$M_Bedeckung,
                   normalized$S_Bedeckung,
                   normalized$D_Windgeschwindigkeit,
                   normalized$M_Windgeschwindigkeit,
                   normalized$S_Windgeschwindigkeit,
                   normalized$Globalstrahlung,
                   normalized$Strompreis)
# Covert matrix to data frame.
columNames <- c("Stunde", "D_Temperatur", "M_Temperatur", "S_Temperatur", "D_Bedeckung", "M_Bedeckung", "S_Bedeckung", "D_Windgeschwindigkeit", "M_Windgeschwindigkeit", "S_Windgeschwindigkeit", "Globalstrahlung","Strompreis")
namePreis = "revPreis"
nameDTemp = "revDTemp"
nameDBed = "revDBed"
nameDWind = "revDWind"
nameSTemp = "revSTemp"
nameSBed = "revSBed"
nameSWind = "revSWind"
nameMTemp = "revMTemp"
nameMBed = "revMBed"
nameMWind = "revMWind"
for(hour in 0:i){
  namePreis = paste("p", namePreis, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,namePreis], type = "0_1"))
  columNames <-c(columNames, namePreis)
  nameGlobalstrahlung = paste("p", nameGlobalstrahlung, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,namePreis], type = "0_1"))
  columNames <-c(columNames, namePreis)
  
  nameDTemp = paste("p", nameDTemp, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameDTemp], type = "0_1"))
  columNames <-c(columNames, nameDTemp)
  nameDBed = paste("p", nameDBed, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameDBed], type = "0_1"))
  columNames <-c(columNames, nameDBed)
  nameDWind = paste("p", nameDWind, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameDWind], type = "0_1"))
  columNames <-c(columNames, nameDWind)
  
  nameSTemp = paste("p", nameSTemp, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameSTemp], type = "0_1"))
  columNames <-c(columNames, nameSTemp)
  nameSBed = paste("p", nameSBed, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameSBed], type = "0_1"))
  columNames <-c(columNames, nameSBed)
  nameSWind = paste("p", nameSWind, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameSWind], type = "0_1"))
  columNames <-c(columNames, nameSWind)
  
  nameMTemp = paste("p", nameMTemp, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameMTemp], type = "0_1"))
  columNames <-c(columNames, nameMTemp)
  nameMBed = paste("p", nameMBed, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameMBed], type = "0_1"))
  columNames <-c(columNames, nameMBed)
  nameMWind = paste("p", nameMWind, sep = "")
  converted <- cbind(converted, normalizeData(strompreise[,nameMWind], type = "0_1"))
  columNames <-c(columNames, nameMWind)
  
}

dataframe <- as.data.frame(converted)
colnames(dataframe) <- columNames

inputs <- dataframe
inputs$Strompreis <- NULL
targets <- dataframe$Strompreis

splitdata <- splitForTrainingAndTest(inputs, targets, ratio=0.1)
#splitdata
```

## Model and analysis
Fit the model. mlp allows to put test data in as well and will test the mlp itself. 
u
Configure the size at your choice.
```{r}
model <- mlp(splitdata$inputsTrain,
             splitdata$targetsTrain,
             size=c(144,72,24),
             maxit=10,
             #hiddenActFunc = "Act_Elliott",
             inputsTest=splitdata$inputsTest,
             targetsTest=splitdata$targetsTest)
model
model.
plotIterativeError(model)
#plotRegressionError(model)
```

## Prediction
## 1h
Taken from prefitted data.
```{r}
prediction <- denormalizeData(model$fittedTestValues,
                              getNormParameters(normalized$Strompreis))
realTest <- denormalizeData(splitdata$targetsTest,
                            getNormParameters(normalized$Strompreis))
result <- cbind(prediction,realTest)
resultdiff <- prediction - realTest
result
plot(result)
plot(resultdiff)

sum((realTest - prediction)^2) / length(realTest)
```

## Continous 24h
```{r}
continous24 <- as.data.frame(splitdata$inputsTest)
continous24result <- NULL
continous24result$real <- denormalizeData(splitdata$targetsTest,
                                    getNormParameters(normalized$Strompreis))
for(i in 1:24)
{
  raw_i <- paste("raw", i)
  pred_i <- paste("pred", i)
  result_i <- paste("result", i)
  diff_i <- paste("diff", i)
  abs_i <- paste("abs", i)
  sum_i <- paste("sum", i)
  continous24result[[pred_i]] <- denormalizeData(predict(model, continous24),
                                        getNormParameters(normalized$Strompreis)) 
  continous24result[[result_i]] <- cbind(continous24result[[pred_i]], continous24result$real)
  continous24result[[diff_i]] <- continous24result$real - continous24result[[pred_i]]
  continous24result[[abs_i]] <- abs(continous24result[[diff_i]])
  continous24result[[sum_i]] <- sum(continous24result[[diff_i]]^2) / length(continous24result[[diff_i]])
  continous24result[[result_i]]
  #plot(continous24result[[pred_i]])
  plot(continous24result[[result_i]])
  #plot(continous24result[[diff_i]])
  #plot(continous24result[[abs_i]])
  
  # Shift previous prices.
  continous24$pppprevPreis <- continous24$ppprevPreis
  continous24$ppprevPreis <- continous24$pprevPreis
  continous24$pprevPreis <- continous24$prevPreis
  # IMPORTANT: denormalize output and normalize input, as different paramters might be used.
  continous24$prevPreis <- continous24result[[result_i]] #normalizeData(continous24result[[pred_i]], getNormParameters(normalized$prevPreis))
}
```
## Single 24h
```{r}
output24 <- NULL
input24 <- splitdata$inputsTest

start <- 0
end <- 120
input24 <- splitdata$inputsTest[start:(end+4),]
for(i in 1:(end-start))
{
  output <- predict(model, input24[i,])
  output24[i-start] <- output
  input24[i+1,"prevPreis"] <- output
  input24[i+2,"pprevPreis"] <- output
  input24[i+3,"ppprevPreis"] <- output
  input24[i+4,"ppprevPreis"] <- output
}

prediction24 <- denormalizeData(output24,
                                getNormParameters(normalized$Strompreis))

realTest24 <- denormalizeData(splitdata$targetsTest[1:(end-start)],
                              getNormParameters(normalized$Strompreis)) 
result24 <- cbind(prediction24,realTest24)
result24diff <- realTest24 - prediction24
result24diffabs <- abs(result24diff)
result24
plot(result24)
plot(result24diff)
plot(result24diffabs)

sum((realTest24 - prediction24)^2) / length(prediction24)
```

Clean up.
```{r message=FALSE}
remove(strompreise)
remove(normalized)
remove(converted)
remove(dataframe)
remove(inputs)
remove(splitdata)
remove(model)
remove(prediction)
remove(realTest)
remove(targets)
remove(result)
```