---
title: "Strompreise"
output: html_notebook
---
Load the libraries.
```{r message=FALSE}
library(RSNNS)
```

Read the data.
```{r}
s <- read.csv("Strompreise.csv")
```

Normalize the data using the RSNNS library.
Definitely omit column DatumUhrzeit as it is the redundant, combined, badly-analyzable value of other columns.
Consider omitting column Jahr as it is the same for every value.
Wochentag is first mapped to values 1-7 using as.numeric and then normalized by RSNNS.
We should optimally convert the data frame to a matrix, normalize it and convert it back to a data frame. But I was not able to do so, so this is the workaround.
```{r}
# Normalize every column.
n = NULL
n$Jahresstunde <- normalizeData(s$Jahresstunde)
# n$Jahr <- normalizeDate(s$Jahr)
n$Monat <- normalizeData(s$Monat)
n$Tag <- normalizeData(s$Tag)
n$Stunde <- normalizeData(s$Stunde)
n$Woche <- normalizeData(s$Woche)
n$Wochentag <- normalizeData(as.numeric(s$Wochentag))
n$Temperatur <- normalizeData(s$Temperatur)
n$Bedeckung <- normalizeData(s$Bedeckung)
n$Windgeschwindigkeit <- normalizeData(s$Windgeschwindigkeit)
n$Globalstrahlung <- normalizeData(s$Globalstrahlung)
n$Strompreis <- normalizeData(s$Strompreis)
# Bind columns to matrix.
c <- cbind(n$Jahresstunde, n$Monat, n$Tag, n$Stunde, n$Woche, n$Wochentag, n$Temperatur, n$Bedeckung, n$Windgeschwindigkeit, n$Globalstrahlung, n$Strompreis)
# Covert matrix to data frame.
d <- as.data.frame(c, col.names=tmp)
# Rename columns.
colnames(d) <- c("Jahresstunde", "Monat", "Tag", "Stunde", "Woche", "Wochentag", "Temperatur", "Bedeckung", "Windgeschwindigkeit", "Globalstrahlung", "Strompreis")
```

Denormalize the column Strompreis of the data frame utilizing the norm parameters of the large list using RSNNS.
```{r}
o <- denormalizeData(d$Strompreis, getNormParameters(n$Strompreis))
```

Print the output.
```{r}
str(o)
# o
```

Clean up.
```{r message=FALSE}
remove(s)
remove(n)
remove(c)
remove(d)
remove(o)
```